---
title: <font size="5"> Main Text Figures</font>
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir="../")
knitr::opts_chunk$set(fig.width=7.5, fig.height=6, fig.path='../figures/fig_data/', warning=FALSE, message=FALSE,cache=FALSE,dev=c('png','pdf'))
```

```{r packages, echo=FALSE}

# packages needed for analytical plots and heatmaps
library(phylosamp)
library(ggplot2)
library(reshape2)
library(dplyr)
library(grid)
library(gridExtra)
library(RColorBrewer)
library(cowplot)
library(viridis)
library(metR)
library(ROCR)
library(R.utils)
library(pbapply)

# addtional packages needed for simulations and plots
library(ape)
#library(outbreaker)
library(dplyr)

# load functions for running simulations
source("source/sim_func.R")
# load extra functions for plotting
source("source/plot_func.R")

# load table containing the density distribution of the mean number of generations between cases
# for different values of R
# the distribution is calculated from 1000 outbreak simulations given R and D, the number of generations
# in these simulations, we use D=floor(ln(R)/ln(1000))
# or the number of generations needed for a final outbreak size of approximately 1000 cases
# to simulate mean distances for other combinations of R and D, see source/simDist.R
mgd <- read.csv("data/sim_distance_dist.csv")

```

<br>

##### **Figure 1.** Sample size and false discovery rate given multiple linkage and multiple transmissions

<br>

##### A
```{r plot_mtml, echo=FALSE, eval=TRUE}
plt.fdr.pairs(chi = 0.999, eta = 0.9, rho.disc = c(0.01,0.05,0.25,0.50,0.75),
       rho.cont = seq(0.01, 1.00, length.out = 100), M.disc = c(50,100,200,500,1000),
       M.cont = seq(2,1000), R = 1, assumption = "mtml")
```

##### B
```{r heatmap_roc_mtml, echo=FALSE, eval=TRUE}
# plot the heatmap and the ROC curve together
# to visualize maximum possible false discovery rate given a mutation rate
chi <- 1-10^(seq(log10(1-.999999), log10(1-0.9), length.out = 100))
eta <- seq(0, 1, length.out = 100)
R0 <- 1
# calculate the density distribution of the mean number of generations between cases given R
mean_gens_pdf <- mgd[mgd["R"]==2][3:dim(mgd)[2]]
plt.heatmap.fdr(chi,eta,rho=0.5,M=200,R=R0,assumption="mtml") +
  geom_point(data=phylosamp::sens_spec_roc(cutoff=1:10,mut_rate=1,mean_gens_pdf=mean_gens_pdf),
             aes(x=specificity,y=sensitivity,fill=NULL,z=NULL),
             col="white", size=4)
```

<br>

##### **Figure 2.** Determining the sensitivity and specificity of a genetic distance cutoff

<br>

```{r gendist_roc_sim, echo=FALSE, fig.width=7.5, fig.height=6, eval=TRUE}

# plot genetic distance distribution from simulated outbreak
# code used to simulate an outbreak
  #sim_data <- true.outbreak(mu = 0.001, R = 2, D = 9, S = 100000, gen = c(0,1), 
  #                          L = 1000, rimport = 0, dimport = 365, C = 70)
  # save(sim_data,file=outbreak)
plt.gendist.sim(outbreak = "data/exampleOutbreak.Rdata")

```

```{r gendist_roc_mut, echo=FALSE, fig.width=7.5, fig.height=6, eval=TRUE}

# calculate the density distribution of the mean number of generations between cases given R
# use R=2 to match simulated outbreak
mean_gens_pdf <- mgd[mgd["R"]==2][3:dim(mgd)[2]]

# make the plot
# use mut_rate = 1 (0.001*genome length) to match simulated outbreak
plt.gendist.mut(mut_rate = 1, mean_gens_pdf = mean_gens_pdf)

```

<br>

##### **Figure 3.** Predicted versus observed false discovery rate in outbreak simulations

<br>

```{r fdr_vs_fdr_sim, echo=FALSE, eval=TRUE}

# first time only: calculate tfdr from simulations and save to file
#calc.tfdr(simdata="data/simdata_var_gen_N10000",rho_values=c(0.1,0.25,0.5,0.75),max_sim_size=2000,
#          sens_spec_method="sim",mgd=mgd,outdir="data/full_data_sim.Rdata")

# plot fdr with sensitivity and specificity calculated from simulations
plt.sim.data(simdata="data/full_data_sim.Rdata",rho_values=c(0.1,0.25,0.5,0.75),max_sim_size=2000,
             filter_by=NA,filter_value=NA,
             xparam="tfdr",yparam="fdr.sub",axis_labels=c("Theoretical FDR","Simulated FDR"))

```

<br>

##### **Table 2.** Bias and error of false discovery rate for simulations with fixed sampling proportion

```{r res_table_sim, echo=FALSE, fig.width=5.5, fig.height=5, eval=TRUE}

# plot the bias from plot using simulated data
bias_table_sim <- fdr.make.tables(saved_data="data/full_data_sim.Rdata",var="bias",
                break_by="tfdr",row_breaks=c(0.0,0.25,0.5,0.75,1.0),
                row_names=c("0-0.25","0.25-0.5","0.5-0.75","0.75-1"))

# plot the error from plot using simulated data
error_table_sim <- fdr.make.tables(saved_data="data/full_data_sim.Rdata",var="err",
                break_by="tfdr",row_breaks=c(0.0,0.25,0.5,0.75,1.0),
                row_names=c("0-0.25","0.25-0.5","0.5-0.75","0.75-1"))

tbias <- textGrob("A (Bias)", gp=gpar(fontsize=12), hjust = 0)
terror <- textGrob("B (Error)", gp=gpar(fontsize=12), hjust = 0)
grid.arrange(tbias,tableGrob(bias_table_sim),terror,tableGrob(error_table_sim),nrow=4,heights=c(1,7,1,7))

```

<br>

##### **Figure 4.** Validation of mutation rate method to calculate sensitivity and specificity

```{r fdr_vs_fdr_mut, echo=FALSE, eval=TRUE}

# first time only: calculate tfdr from simulations and save to file
#calc.tfdr(simdata="data/simdata_var_gen_N10000",rho_values=c(0.1,0.25,0.5,0.75),max_sim_size=2000,
#          sens_spec_method="mutrate",mgd=mgd,outdir="data/full_data_mutrate.Rdata")

# plot fdr with sensitivity and specificity calculated from simulations
plt.sim.data(simdata="data/full_data_mutrate.Rdata",rho_values=c(0.1,0.25,0.5,0.75),max_sim_size=2000,
             xparam="tfdr",yparam="fdr.sub",axis_labels=c("Theoretical FDR","Simulated FDR"))

```

<br>

##### **Figure 5.** Distribution of raw errors using mutation rate method

```{r error_hist_all, echo=FALSE, fig.width=7.5, fig.height=4, eval=TRUE, warning=FALSE}

err_hists <- plot.error.hist(saved_data="data/full_data_mutrate.Rdata",single_value=FALSE)

blank <- textGrob("",gp=gpar(fontsize=8))
rho10 <- textGrob("Sampling proportion: 0.1",gp=gpar(fontsize=8),hjust=0.37)
rho25 <- textGrob("Sampling proportion: 0.25",gp=gpar(fontsize=8),hjust=0.37)
rho50 <- textGrob("Sampling proportion: 0.5",gp=gpar(fontsize=8),hjust=0.37)
rho75 <- textGrob("Sampling proportion: 0.75",gp=gpar(fontsize=8),hjust=0.37)
toplabel <- textGrob("FDR", gp=gpar(fontsize=8), hjust = 0.3)
midlabel <- textGrob("Sensitivity", gp=gpar(fontsize=8), hjust = 0.4)
botlabel <- textGrob("Specificity", gp=gpar(fontsize=8), hjust = 0.4)
grid.arrange(blank,rho10,rho25,rho50,rho75,
             toplabel,err_hists[[1]],err_hists[[2]],err_hists[[3]],err_hists[[4]],
             midlabel,err_hists[[5]],err_hists[[6]],err_hists[[7]],err_hists[[8]],
             botlabel,err_hists[[9]],err_hists[[10]],err_hists[[11]],err_hists[[12]],
             nrow=4,widths=c(1.1,3,3,3,3),heights=c(1,8,8,8))

```

<br>

##### **Figure 6.** Effect of different generation distributions on specificity

```{r sens_spec_main, echo=FALSE, eval=TRUE, fig.show = 'hide'}

# first time only: calculate tfdr from simulations and save to file

# load generation distribution data
rho_values=c(0.1,0.25,0.5,0.75)
mgd_list <- vector(mode = "list", length = 4)
for (i in 1:4){
  simpath <- paste("data/gendata_var_sim_N10000","_rho",as.character(100*rho_values[i]),".csv",sep="")
  mgd_list[[i]] <- read.csv(simpath)
  }

#calc.tfdr(simdata="data/simdata_var_gen_N10000",rho_values=rho_values,max_sim_size=2000,
#          sens_spec_method="mutrate",mgd=mgd_list,mgd_type="sim",outdir="data/full_data_mutrate_gens.Rdata")

# plot specificity from mutrate data
p_mutrate_spec <- plt.sim.data(simdata="data/full_data_mutrate.Rdata",rho_values=0.75,max_sim_size=2000,
             xparam="t.chi",yparam="chi.full",include_titles=FALSE,
             axis_labels=c("Theoretical specificity","Simulated specificity"))

# plot specificity from gens data
p_gens_spec <- plt.sim.data(simdata="data/full_data_mutrate_gens.Rdata",rho_values=0.75,max_sim_size=2000,
             xparam="t.chi",yparam="chi.full",include_titles=FALSE,
             axis_labels=c("Theoretical specificity","Simulated specificity"))

```

```{r sens_spec_main_plot, echo=FALSE, fig.width=7.5, fig.height=3, eval=TRUE}

# arrange the plots as desired for this figure
Alabel <- textGrob("A", gp=gpar(fontsize=10,fontface="bold"),hjust=-0.1)
Blabel <- textGrob("B", gp=gpar(fontsize=10,fontface="bold"),hjust=-0.1)
lefttitle <- textGrob("Generation distribution: average", gp=gpar(fontsize=10,fontface="bold"),hjust=0.55)
righttitle <- textGrob("Generation distribution: actual", gp=gpar(fontsize=10,fontface="bold"),hjust=0.55)
titles <- arrangeGrob(Alabel,lefttitle,Blabel,righttitle,ncol=4,widths=c(1,18,1,18))
plots <- arrangeGrob(p_mutrate_spec,p_gens_spec,ncol=2)
grid.arrange(titles,plots,nrow=2,heights=c(1,6))

```


```{r close_plots, echo=FALSE, results='hide'}

dev.off()

```